---
title: "Statues Named John: London Memorials"
format:
  dashboard:
    orientation: rows
    # Rows keep their natural height and the page scrolls, instead of every
    # row being squeezed into the window (charts shrank to thumbnails)
    scrolling: true
    theme: cosmo
    include-in-header:
      - text: |
          <style>
          /* Tab content scrolls instead of being squashed */
          .tab-pane {
            overflow-y: auto !important;
            max-height: 80vh;
          }
          .tab-pane .cell-output-display {
            overflow: auto !important;
          }
          /* Each page grid splits the window height between its rows, so on
             short windows charts became thumbnails and tables lost their rows.
             Give every row a floor and let it grow; the page scrolls instead. */
          .dashboard-scrolling .dashboard-page > .bslib-grid {
            grid-template-rows: none !important;
            grid-auto-rows: minmax(600px, auto) !important;
          }
          .dashboard-scrolling .cell-output-display img {
            min-height: 380px;
            object-fit: contain;
          }
          .dashboard-scrolling .card-body > table,
          .dashboard-scrolling .cell-output-display table {
            flex-shrink: 0;
          }
          /* Click-to-zoom fullscreen overlay for maps and charts */
          .fullscreen-overlay {
            position: fixed !important;
            top: 0 !important;
            left: 0 !important;
            width: 100vw !important;
            height: 100vh !important;
            z-index: 9999 !important;
            background: #ffffff !important;
            padding: 20px !important;
            overflow: auto !important;
            cursor: zoom-out !important;
            box-sizing: border-box !important;
          }
          .fullscreen-overlay .leaflet {
            height: calc(100vh - 80px) !important;
          }
          .fullscreen-overlay img {
            max-height: calc(100vh - 80px);
            width: auto;
          }
          .card:has(.leaflet),
          .card:has(.cell-output-display img) {
            cursor: zoom-in;
          }
          .cell-output-display .bslib-full-screen-enter {
            display: none !important;
          }
          </style>
          <script>
          document.addEventListener('click', function(e) {
            var overlay = e.target.closest('.fullscreen-overlay');
            if (overlay) {
              // Leaflet keeps handling clicks inside the map itself
              if (e.target.closest('.leaflet')) return;
              overlay.classList.remove('fullscreen-overlay');
              window.dispatchEvent(new Event('resize'));
              return;
            }
            if (e.target.closest('.leaflet')) return;
            if (e.target.closest('.bslib-full-screen-enter')) return;
            if (e.target.closest('.nav-link, details, summary, a')) return;
            var card = e.target.closest('.card');
            if (!card) return;
            if (!card.querySelector('.leaflet, .cell-output-display img')) return;
            card.classList.add('fullscreen-overlay');
            setTimeout(function() { window.dispatchEvent(new Event('resize')); }, 100);
          });
          </script>
execute:
  echo: false
  warning: false
  message: false
vignette: >
  %\VignetteIndexEntry{Comparing London Memorials: Johns, Women, and Dogs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
library(dplyr)
library(ggplot2)
library(sf)
library(targets)
library(leaflet)

# All results come from the targets pipeline; run targets::tar_make() first.
# Absolute store path: Quarto may execute this from inst/qmd/.
root <- rprojroot::find_root(rprojroot::is_r_package)
store_path <- file.path(root, "_targets")
if (dir.exists(store_path)) {
  all_memorials <- tar_read(all_memorials, store = store_path)
  summary_table <- tar_read(summary_table, store = store_path)
  category_plot <- tar_read(category_plot, store = store_path)
  memorial_map_plot <- tar_read(memorial_map_plot, store = store_path)
  memorial_interactive_map <- tar_read(memorial_interactive_map, store = store_path)
  johns_comparison <- tar_read(johns_comparison, store = store_path)
  gender_analysis <- tar_read(gender_analysis, store = store_path)
  source_status <- tar_read(source_status, store = store_path)
  # Code behind each target, read from the plan files (#81)
  target_code <- tar_read(target_code, store = store_path)
} else {
  message("Targets pipeline store not found. Run targets::tar_make() first.")
  all_memorials <- data.frame()
  summary_table <- data.frame()
  category_plot <- ggplot()
  memorial_map_plot <- ggplot()
  memorial_interactive_map <- leaflet()
  johns_comparison <- list(message = "Pipeline not run", claim_validated = FALSE,
                           total_statues = NA_real_, john_statues = 0,
                           john_percent_label = "n/a", woman_statues = 0,
                           woman_only_statues = 0, mixed_statues = 0,
                           woman_percent_label = "n/a", unknown_statues = 0,
                           unknown_percent_label = "n/a",
                           classification_threshold = "n/a",
                           wikidata_confidence = c(exact = "n/a", partial = "n/a"))
  gender_analysis <- list(top_names_by_gender = data.frame())
  source_status <- data.frame()
  target_code <- list()
}

pkg_version <- read.dcf(file.path(root, "DESCRIPTION"), fields = "Version")[1, 1]
generated_time <- format(Sys.time(), "%Y-%m-%d %H:%M %Z")
n_mapped <- sum(!is.na(all_memorials$lat) & !is.na(all_memorials$lon))

top_names <- function(gender) {
  gender_analysis$top_names_by_gender %>%
    dplyr::filter(inferred_gender == gender) %>%
    dplyr::select(Name = extracted_names, Count = n, Percent = percent_label) %>%
    dplyr::arrange(dplyr::desc(Count))
}
top_f <- top_names("Female")
top_m <- top_names("Male")
john_rank <- match("John", top_m$Name)
by_source <- if (nrow(all_memorials) > 0) table(all_memorials$source_name) else table(character(0))
```

# Overview

## Row

### Headline {.tabset width=40%}

#### Finding

**Are there more statues of men named John than of women in London?**
In this dataset of `r johns_comparison$total_statues` memorials: **`r johns_comparison$john_statues` Johns (`r johns_comparison$john_percent_label`)** vs **`r johns_comparison$woman_statues` statues of women (`r johns_comparison$woman_percent_label`)**.
The claim is **`r if (isTRUE(johns_comparison$claim_validated)) "supported" else "refuted"`** for London.
`r johns_comparison$unknown_statues` memorials (`r johns_comparison$unknown_percent_label`) have no gender that clears the confidence threshold.

#### Limerick

| In England's grand capital, dear,
| More Johns than dames, it appears.
| For each woman of might,
| A John's in plain sight,
| Confirming our gendered-statue fears.

### Gender Summary {.tabset width=60%}

#### Table

::: {.card title="Memorials by inferred gender"}

```{r}
knitr::kable(summary_table, align = "lrr")
```

**Caption:** How are London's memorials split by the gender of who they commemorate?
One row per category, from `r johns_comparison$total_statues` memorials combined from Wikidata, OpenStreetMap and Historic England after de-duplication.
"Mixed" statues show a woman together with a man; the headline counts them as statues of women (`r johns_comparison$woman_only_statues` Female + `r johns_comparison$mixed_statues` Mixed = `r johns_comparison$woman_statues`).
"Unknown" means no gender reached the confidence threshold (see Reference).

:::

#### Code

```{r}
#| results: asis
cat(target_code[["summary_table"]])
```

## Row

### Representation {.tabset}

#### Chart

::: {.card title="Memorials by inferred gender"}

```{r}
#| fig-width: 9
#| fig-height: 5
category_plot
```

**Caption:** How large is the gap between statues of men and of women?
Bar heights are counts of memorials per inferred gender, labelled with count and share.
Statues of women, including mixed ones, are `r johns_comparison$woman_percent_label` of the total.
The Unknown bar (`r johns_comparison$unknown_percent_label`) is the main uncertainty: if many of those are women, the gap narrows.

:::

#### Code

```{r}
#| results: asis
cat(target_code[["category_plot"]])
```

# Top Names

## Row

### Most Common First Names {.tabset}

#### Female

::: {.card title="Most common first names: women"}

```{r}
knitr::kable(top_f, align = "lrr")
```

**Caption:** Which first names recur among statues of women?
Counts are memorials whose subject's first name is the given name, among those classified Female; percentages are of all statues of women.
The most common is `r top_f$Name[1]` (`r top_f$Count[1]`).
The list comes from the name rules, so a word that is not a name can slip in.

:::

#### Male

::: {.card title="Most common first names: men"}

```{r}
knitr::kable(top_m, align = "lrr")
```

**Caption:** Is John really the most common name among statues of men?
Counts are memorials whose subject's first name is the given name, among those classified Male; percentages are of all statues of men.
John ranks `r if (is.na(john_rank)) "outside this list" else paste0("number ", john_rank, " (", top_m$Count[john_rank], ")")`.
This count is of the name John only; the Overview headline also counts Jon and Jean.
The list comes from the name rules, so a word that is not a name can slip in.

:::

#### Code

```{r}
#| results: asis
cat(target_code[["gender_analysis"]])
```

# Map

## Row

### Memorial Locations {.tabset}

#### Interactive

::: {.card title="London memorials (click a cluster to zoom; click the card to enlarge)"}

```{r}
#| out-height: 550px
memorial_interactive_map
```

**Caption:** Where in London are the memorials, and which source records each one?
`r n_mapped` of `r nrow(all_memorials)` memorials have coordinates; clusters show counts and split as you zoom in.
Click a marker for that memorial's details.

:::

#### By source

::: {.card title="Memorials in London by source"}

```{r}
#| fig-width: 9
#| fig-height: 7
memorial_map_plot
```

**Caption:** Do the three sources cover the same parts of London?
Each point is one memorial, coloured by the source whose record was kept after de-duplication.
Counts by source: `r paste(names(by_source), by_source, sep = " ", collapse = ", ")`.

:::

#### Code

```{r}
#| results: asis
cat(target_code[["memorial_interactive_map"]])
cat(target_code[["memorial_map_plot"]])
```

# Johns vs Women

## Row

### The Claim {.tabset}

#### Finding

The campaign [Statues for Equality](https://statuesforequality.com/pages/london) claims that:

- "The percentage of women's statues in the UK that aren't mythical or royal is approximately 3%"
- "More statues of men named John dotted around the country than of women!"

Our London data: **`r johns_comparison$message`**

**Validation:** the claim that there are "more statues of men named John than of women" is **`r if (isTRUE(johns_comparison$claim_validated)) "supported" else "refuted"`** by this London dataset.

- **Men named John:** `r johns_comparison$john_statues` (`r johns_comparison$john_percent_label`)
- **Women:** `r johns_comparison$woman_statues` (`r johns_comparison$woman_percent_label`), of which `r johns_comparison$woman_only_statues` depict only women and `r johns_comparison$mixed_statues` a woman together with a man

Percentages are of all `r johns_comparison$total_statues` memorials in the dataset.
The campaign's claim is UK-wide; this dataset covers London only.

#### Overlap

The two counts can overlap.
A statue of a John together with a woman, such as "Katherine Mansfield and John Middleton Murry", counts once as a John statue and once as a statue of a woman.

#### Code

```{r}
#| results: asis
cat(target_code[["johns_comparison"]])
```

# Reference

## Row

### Reference Details {.tabset}

#### How Gender Is Decided

Many records name the object, not the person: "Statue of Sir John Millais in North Forecourt of Tate Gallery".
We first read the subject from that text (here "Sir John Millais").

If first names and titles cannot decide it, we ask Wikidata whether the subject is a person and, if so, their sex.
An exact name match scores `r johns_comparison$wikidata_confidence[["exact"]]` and a partial match `r johns_comparison$wikidata_confidence[["partial"]]`.

A gender is used only when its confidence is at least the classification threshold, currently `r johns_comparison$classification_threshold`.
The same threshold applies to first-name predictions.
It is set in one place, `inst/extdata/params.csv`; setting it to 1 keeps only certain answers, and the rest stay Unknown.

#### Data Sources

::: {.card title="Rows fetched per source"}

```{r}
knitr::kable(source_status, align = "lrrl")
```

**Caption:** Did every data source respond, and how many rows did each return?
`ok` means the fetch succeeded; an unavailable source is recorded with its reason rather than counted as empty.
`wikidata_people` is the person lookup used to decide gender for subjects the name rules cannot (see How Gender Is Decided).

- [Wikidata](https://www.wikidata.org/): statues and public art in London, via SPARQL.
- [OpenStreetMap](https://www.openstreetmap.org/): memorials and statues, via the Overpass API.
- [Historic England / GLHER](https://historicengland.org.uk/): heritage records.
- Code: [statuesnamedjohn on GitHub](https://github.com/JohnGavin/statues_named_john), pipeline in `R/tar_plans/`.

:::

#### About

This dashboard compares memorials in London honouring three groups:

1. **Men named John**: how common one male first name is among commemorated figures.
2. **Women** of any name: gender representation in public memorials.
3. **Dogs**, and animals generally: a quirky baseline.

Every number on these pages is computed by the `targets` pipeline; the **Code** tab on each card shows the pipeline step that produced it.

#### Session Info

```{r}
sessionInfo()
```

---

*Dashboard generated `r generated_time` using statuesnamedjohn v`r pkg_version`.*

::: {.callout-note appearance="minimal"}
**Disclaimer:** `statuesnamedjohn` is an independent open-source project ([website](https://johngavin.github.io/statues_named_john/), [GitHub](https://github.com/JohnGavin/statues_named_john)). It is not affiliated with or endorsed by [Statues for Equality](https://statuesforequality.com/).
:::
